---
title: "data_comparison"
author: "Hannah Dempsey"
output: github_document
---

# Comparing Field Site Statistics

Set up:

```{r}
#Week 3 (loading in packages and read_csv)

#install.packages("tidyverse")
library(tidyverse)

#creating function to plot ecological data facet wrapped by species
#Week 11 (creating functions)
ecol_per_species <- function(data, species_vector, y_axis, y_label) {
  graph <- data %>% 
    filter(species %in% species_vector) %>% 
    ggplot(., aes(x = year, y = {{y_axis}}, color = site)) +
    geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    facet_wrap(~species, scales = "free_y") + 
    labs(y = y_label, x = "Year") +
    theme_bw()
  return(graph)
}
```

### Biomass Data

#### By species:

Adding site column to dataframes:

```{r}
#Week 3 (read_csv), Week 4 (mutate), and Week 9 (file paths)
#loading in the data
saguaro_biomass <- read_csv("../data_clean/saguaro_biomass.csv")
orpi_biomass <- read_csv("../data_clean/orpi_biomass.csv")

#adding site column to each dataframe
saguaro_biomass <- saguaro_biomass %>% 
  mutate(site = "Saguaro National Park")
saguaro_biomass

orpi_biomass <- orpi_biomass %>% 
   mutate(site = "Organ Pipe National Monument")
orpi_biomass
```

Bind dataframes into one:

```{r}
#Week 4 (bind_rows)
all_biomass <- bind_rows(saguaro_biomass, orpi_biomass)
all_biomass
```

Plotting annual biomass of each species:

```{r}
#Week 2 (vectors)

#creating vector of species present in both sites
shared_species <- c("albigula", "amplus", "baileyi", "eremicus", "merriami", "pennicillatus", "torridus")

#plotting data
#Week 5 (data visualization)
ecol_per_species(all_biomass, shared_species, biomass_g, "Biomass (g)")
ggsave("../figures/all_biomass.png")
```

#### By community:

Adding site column to all dataframes:

```{r}
#Week 3 (read_csv), Week 4 (mutate), and Week 9 (file paths)

#loading in the data
saguaro_total_biomass <- read_csv("../data_clean/saguaro_total_biomass.csv")
orpi_total_biomass <- read_csv("../data_clean/orpi_total_biomass.csv")

#adding site column to each dataframe
saguaro_total_biomass <- saguaro_total_biomass %>% 
  mutate(site = "Saguaro National Park")
saguaro_total_biomass

orpi_total_biomass <- orpi_total_biomass %>% 
   mutate(site = "Organ Pipe National Monument")
orpi_total_biomass
```

Bind dataframes into one:

```{r}
#Week 4 (bind_rows)

all_total_biomass <- bind_rows(saguaro_total_biomass, orpi_total_biomass)
all_total_biomass
```

Plotting community biomass:

```{r}
#Week 5 (data visualization)

ggplot(all_total_biomass, aes(x = year, y = biomass_g, color = site)) +
  geom_point() +
  geom_line() +
  labs(y = "Total Biomass (g)", x = "Year") +
  theme_bw()
ggsave("../figures/orpi_and_saguaro_biomass.png")
```

### Abundance Data

Loading in datasets:

```{r}
#Week 3 (read_csv and select), Week 4 (mutate), and Week 9 (file paths)

saguaro_abund <- read_csv("../data_clean/saguaro_abund_trapnights.csv")
orpi_abund <- read_csv("../data_clean/orpi_abund_trapnights.csv")

#removing superfluous columns
saguaro_abund <- saguaro_abund %>% 
  select(1, 2, 6)
saguaro_abund

orpi_abund <- orpi_abund %>% 
  select(1, 2, 5)
orpi_abund

#adding site column to each dataframe
saguaro_abund <- saguaro_abund %>% 
  mutate(site = "Saguaro National Park")
saguaro_abund

orpi_abund <- orpi_abund %>% 
   mutate(site = "Organ Pipe National Monument")
orpi_abund
```

Bind dataframes into one

```{r}
#Week 4 (bind_rows)

all_abund <- bind_rows(saguaro_abund, orpi_abund)
all_abund
```

Plotting abundance of each species:

```{r}
#Week 5 (data visualization)

ecol_per_species(all_abund, shared_species, abund_per_trapnight, "Abundance per Trapnight")
ggsave("../figures/all_abund.png")
```
